---
title: "Aides Drug"
author: "Emann Malik"
date: "March 20, 2021"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```
```{r echo=FALSE}
triUmp$treatment <- as.factor(triUmp$treatment)
triUmp$infected <- as.factor(triUmp$infected)
triUmp$LGBTQ <- as.factor(triUmp$LGBTQ)
triUmp$DrugUser <- as.factor(triUmp$DrugUser)
triUmp$gender <- as.factor(triUmp$gender)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->
  
A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.
  
In this report, we will be trying to show whether or not triUmpsal, a new drug that is intended to prevent the contraction of AIDS, is effective in females. The data we will be using comes from the triUmp data set, focusing mainly on the factor variables "treatment," "infected," and "gender" (the "F" level). We will be attempting to answer this research question: Is triUmpsal effective in preventing females from contracting AIDS?
  
# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
  
We will be conducting a variable analysis via the graphical method of a bar plot and the numerical method of a summary (mean, minimum, standard deviation, etc.) of the data. We will need the data manipulation tool called "filter" from the package dplyr to retrieve only the females from the "gender" variable and see how the drug has worked for them. The inferential tools of the chi-squared and Fisher's exact test will also be utilized to help finalize our answer to the research question.
  
```{r}
females <- filter(triUmp, gender == "F")
```
  
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
  
```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```
  
The bar plot above shows that, of all the females included in this clinical trial who got infected with AIDS, about 60% had received the triUmpsal drug and just over 40% had received the placebo. In other words, the drug seems to be increasing the likelihood that a female will get infected with the disease!
  
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
  
```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```
  
The first table above shows that much more than half - 58.82% - of the females who became infected with AIDS took the triUmpsal drug, a number that surpasses the 41.18% of infected females who took the placebo. These amounts are reflected in the second table, which shows that more infections occurred in females who took the drug - 3.69% - than in females who took the placebo - 2.42%. Once again, we are seeing evidence that shows the drug is much less effective than it should be.
  
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
